### 2019-11-12 PaperDraft.Rmd
echo = FALSE

---
title: "Draft"
author: " "
date: "November 11, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### Motivation

- Explination of Game, Set, Match

### Methods

- Where we got our data

- Probability of winning a game, set, match

### Reproducability

### Extension