### 2019-12-11 Group1Prez.Rmd
New Tourney Data

---
title: "Group 1: Tennis "
author: "Olivia Beck, Emma Lewis, Ryan Volkert"
date: "December 12, 2019"
output: beamer_presentation
header-includes: 
  - \usepackage{booktabs} 
  - \usepackage{longtable} 
  - \usepackage{array} 
  - \usepackage{multirow} 
  - \usepackage{wrapfig} 
  - \usepackage{float} 
  - \usepackage{colortbl} 
  - \usepackage{pdflscape} 
  - \usepackage{tabu} 
  - \usepackage{threeparttable} 
  - \usepackage{threeparttablex} 
  - \usepackage[normalem]{ulem} 
  - \usepackage{makecell} 
  - \usepackage{xcolor}

---

```{r setup, include=FALSE, warning=FALSE}
source("funs.R")
library(knitr)
options(kableExtra.latex.load_packages = FALSE) 
library(kableExtra)
library(tidyverse)
set.seed(400)
```

## Project

- Project based off a paper by Paul K. Newton and Joseph B. Keller called Probability of Winning at Tennis 1. Theory and Data

- Paper discussed how the probability of winning a game, a set, and a match in tennis are computed based on each player's probability of winning a point on a serve

- Based on other papers the conclusion was made to treat serves in tennis as independent identically distributed (iid) random variables

- It is noted that points in tennis are not iid for most purposes this is not a bad assumption as the divergence from iid is small

- Both two out of three and three out of five set matches were considered aloowing for 13-point tiebreaker in each set if needed

## The Game of Tennis

- Point: Smallest unit of measurement (Love-15-30-40-game)

- Game: A game is one when a player reaches 4 points with at least a 2 point advantage

- Set: A set consists of 6 games and is won by the player who reaches 6 games first

- Advantage Set: If a game score of 6-6 is reached and advantage set rules are used, a player can only win a set with a 2 game lead

- Matches: Best of 3 set (for women) or 5 sets (for men)

- Tie-break game: If a game score of 6-6 is reached and tie-break set rules are used. In a tie-break game, a player/team must reach 7 points with a two point advantage to win

## Game of Tennis

```{r,echo=FALSE, fig.align='center'}
include_graphics("Scoring.jpg")
```

## Values

Things we need to know:

- Empirical probability of winning a rally on serve

- Empirical probability of winning a game on serve  \pause

Things we need to calculate:

- Theoretical probability of winning a game on serve

- Theoretical probability of winning a set

- Theoretical probability of winning the tie-breaker

- Theoretical probability of winning a match

## Theoretical Probabilities: Rally and Game

$$\text{P(A Winning a Rally | A Served)} = p_A^R = \frac{\text{Points Won On Serve}}{\text{Points Served}}$$ \pause

$$\text{P(A Winning a Game | A Served)}$$
$$=p_A^G = (p_A^R)^4[1+4q_A^R+10(q_A^R)^2] +20(p_A^Rq_A^R)^3(p_A^R)^2[1-2p_A^Rq_A^R]^{-1}$$

- Where Probability A loses a rally given A served: 
$$q_A^R = 1-\frac{\text{Points Won On Serve}}{\text{Points Served}}$$

## Theoretical Probabilities: Set

$$ \text{P(A Winnning a Set | A Served)} $$
$$= p_A^S =\sum_{j=0}^4p_A^S(6,j) + p_A^S(7,5) + p_A^S(6,6)p_A^T$$ \pause

Where, $p_A^S(i,j)$ is defined recursivly as:

$p_A^S(0,0) = 1$, $p_A^S(i,j) = 0$ if i<0 or j<0.

- if i+j -1 is even: $p_A^S(i,j) = p_A^S(i-1,j)p_A^G + p_A^S(i, j-1)q_A^G$

    - omit i-1 term if j=6 and i<6

    - omit j-1 term if i=6 and j<6

- if i+j -1 is odd: $p_A^S(i,j) = p_A^S(i-1,j)q_B^G + p_A^S(i, j-1)p_B^G$

    - omit i-1 term if j=6 and i<6

    - omit j-1 term if i=6 and j<6

## Theoretical Probabilities: Tie Breaker

$$\text{P(A Winning the Tie Breaker | A Served Initially)} $$
$$=p_A^T = \sum_{j=0}^5p_A^T(7,j) +  p_A^T(6,6)\sum_{n=1}^{\infty} p_A^T(n+2, n) $$ \pause

Where,  $p_A^T(i,j)$ is defined recursivly as:

$p_A^T(0,0) = 1$, $p_A^T(i,j) = 0$ if i<0 or j<0.

- if $i+j -1 \mod 4 \equiv 0 \; \text{or} \;3$: $p_A^S(i,j) = p_A^T(i-1,j)p_A^R + p_A^T(i, j-1)q_A^R$

    - omit i-1 term if j=7 and i<7

    - omit j-1 term if i=7 and j<7

- if $i+j -1 \mod 4 \equiv 1 \; \text{or} \; 2$: $p_A^T(i,j) = p_A^T(i-1,j)q_B^R + p_A^T(i, j-1)p_B^R$

    - omit i-1 term if j=7 and i<7

    - omit j-1 term if i=7 and j<7

## Theoretical Probabilities: Match

Women's (Best of 3):
$$\text{P(A Winning the Match | A Served First)}$$
$$=p_A^M = (p_A^S)^2 + 2(p_A^S)^2p_B^S$$ \pause

Men's (Best of 5):
$$\text{P(A Winning the Match | A Served First)}$$ 
$$=p_A^M = (p_A^S)^3 + 3(p_A^S)^3p_B^S + 6(p_A^S)^3(p_B^S)^2$$

Where:

- $p_A^S=\text{Probabilty A wins a set given A served}$

- $p_B^S=\text{Probabilty B wins a set given B served}$

## Data being used 
Data used in paper:

  - Men and Women semifinalists in the 2002 U.S Open tournament

  - Men and Women semifinalists in the 2002 Wimbledon tournament

Data used in our simulation:

  - Men and Women semifinalists in the 2019 Wibledon tournament

## Results of Game 
```{r, echo = FALSE}

TDWomen <- as.data.frame(matrix(0, nrow = 4, ncol = 3))

colnames(TDWomen) <- c("P(Win a Rally)", "Empirical P(Win a Game)", "Paper P(Win a Game)" )

TDWomen[1, ] <- c(0.69, 0.71, 0.89)
TDWomen[2, ] <- c(0.63, 0.80, 0.79)
TDWomen[3, ] <- c(0.65, 0.85, 0.83)
TDWomen[4, ] <- c(0.63, 0.77, 0.79)

OurDatWomen <- as.matrix(round(c(pg(TDWomen[1,1]), pg(TDWomen[2,1]), pg(TDWomen[3,1]), pg(TDWomen[4,1])),2))

TDWomen$`Our P(Win a Game)` <-(OurDatWomen)

rownames(TDWomen) <- c("S. Williams", "V. Williams", "L. Davenport", "A. Mauresmo")

kable(TDWomen, format = "latex", booktabs = TRUE, caption = "Data for the Womens Semifinalists in the 2002 U.S. Open Tournament
")%>%
          kable_styling(latex_options = "scale_down")

```
\pause
```{r, echo = FALSE}

TDMen<- as.data.frame(matrix(0, nrow = 4, ncol = 3))

colnames(TDMen) <- c("P(Win a Rally)", "Empirical P(Win a Game)", "Paper P(Win a Game)" )

TDMen[1, ] <- c(0.73, 0.95, 0.93)
TDMen[2, ] <- c(0.66, 0.87, 0.85)
TDMen[3, ] <- c(0.67, 0.85, 0.86)
TDMen[4, ] <- c(0.68, 0.90, 0.88)

OurDatWMen <- as.matrix(round(c(pg(TDMen[1,1]), pg(TDMen[2,1]), pg(TDMen[3,1]), pg(TDMen[4,1])),2))

TDMen$`Our P(Win a Game)` <-(OurDatWMen)

rownames(TDMen) <- c("P. Sampras", "A. Agassi", "L. Hewitt", "S. Schalken")

kable(TDMen, format = "latex", booktabs = TRUE, caption = "Data for the Mens Semifinalists in the 2002 U.S. Open Tournament
")%>%
          kable_styling(latex_options = "scale_down")

```

## Tournament Results: Women
```{r, echo=FALSE}
Tourney1_Women = data.frame(`P(Winning a Rally)` = TDWomen[,1],
                            `Paper P(Winning the Tournement)` = c(.2872, .3599, .0482, .3074))
#`Our P(Winning the Tournement )` =round( pTC(TDWomen[1,1],TDWomen[2,1],TDWomen[3,1],TDWomen[4,1], 2), 4)

rownames(Tourney1_Women)<- c("S. Williams","L. Davenport",  "A. Mauresmo", "V. Williams")

T1<- table(base::sample(rownames(Tourney1_Women), 2000, prob = Tourney1_Women$Paper.P.Winning.the.Tournement., replace = TRUE))

Winner1 <- which(T1 == max(T1))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") + 
  ylab("P(Winning the Tournement") +
  ggtitle("Which Player Won the Tournement?")

```

## Tournament Results: Men
```{r, echo=FALSE}
Tourney1_Men = data.frame(`P(Winning a Rally)` = TDMen[,1],
                            `Paper P(Winning the Tournement)` = c(.6747, .0851, .0945, .1457))
#`Our P(Winning the Tournement )` =round( pTC(TDMen[1,1],TDMen[2,1],TDMen[3,1],TDMen[4,1], 2), 4)

rownames(Tourney1_Men)<- c("P. Sampras","S. Schalken",  "A. Agassi", "L. Hewitt")

T1<- table(base::sample(rownames(Tourney1_Men), 2000, prob = Tourney1_Men$Paper.P.Winning.the.Tournement., replace = TRUE))

Winner1 <- which(T1 == max(T1))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") +
  ylab("P(Win the Tournement)") +
  ggtitle("Which Player Won the Tournement?")

```

## 2019 Wimbledon: Women

```{r, warning=FALSE, echo=FALSE}
WimWomen <- read_csv("2019 Womens Wimbledon.csv")

WimWomen$`P(Winning a Rally)` <- WimWomen$`Points Won on Serve`/WimWomen$`Total Points Served`

WimWomen <- as.data.frame(WimWomen)

WimWomen$`P(Winning Tournement)`<- pTC(WimWomen[1,4][1], WimWomen[2,4][1],WimWomen[3,4][1],WimWomen[3,4][1],2)

T1<- table(base::sample(WimWomen$Name, 2000, prob = WimWomen$`P(Winning a Rally)`, replace = TRUE))

Winner1 <- which(T1 == max(T1))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") +
  ylab("P(Win the Tournement)")
  ggtitle("Which Player Won the Tournement?")

```

## 2019 Wimbledon: Men

```{r, warning=FALSE, echo=FALSE}
WimMen <- read_csv("2019 Men's Wimbledon.csv")

WimMen <- as.data.frame(WimMen)

WimMen$`P(Winning a Rally)` <- WimMen$`Points Won on Serve`/ WimMen$`Total Points on Serve`

WimMen$`P(Winning Tournement)`<- pTC(WimMen[1,4][1], WimMen[2,4][1],WimMen[3,4][1],WimMen[3,4][1],2)

T1<- table(base::sample(WimMen$Name, 2000, prob = WimMen$`P(Winning a Rally)`, replace = TRUE))

Winner1 <- which(T1 == max(T1))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab("Players") +
  ggtitle("Which Player Won the Tournement?")

```